Bayesian RKHS-based methods in functional regression\support{Support information of the article.}

### 2022-03-05 paper/paper.tex
Some corrections

The most common linear regression model is the classical functional linear regression model, originally introduced in the first edition of the book by~\citet{ramsay2005functional} \incomment{fue realmente así?}. It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^n\) for that of the functional space \(L^2[0,1]\):
\begin{equation}\label{eq:l2-linear-model}
Y = \beta_0 + \dotprod{X}{\beta} + \epsilon = \beta_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
\end{equation}
\incomment{mejor poner el producto con el subíndice \(\langle \cdot , \cdot \rangle_2\)?}~where \(\beta_0\in \R\) and \(\epsilon\) is an independent error term with certain simplifying assumptions. In this case, the main parameter \(\beta=\beta(t)\) is a member of the infinite-dimensional space \(L^2[0, 1]\). Aside from the fact that functional (i.e. infinite-dimensional) optimization is a hard problem in and of itself \incomment{cita?}, there are a number of more subtle drawbacks that arise in this \(L^2\)-framework. For instance, \(L^2[0, 1]\) is a big space that contains many non-smooth or ill-behaved functions, but incidentally the model given by~\eqref{eq:l2-linear-model} does not include a finite-dimensional model based on a linear combination of projections of \(X\) as a particular case \citep[see][]{berrendero2019rkhs}. Moreover, the non-invertibility of the covariance operator (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.

A similar \(L^2\)-based logistic regression model can be derived for the binary classification problem via the logistic function:
\begin{equation}\label{eq:l2-logistic-model}
  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\beta_0 - \dotprod{X}{\beta}\}},
\end{equation}
where \(\beta_0 \in \R\) and \(\beta \in L^2[0, 1]\). In this case, the most common way of estimating the slope function \(\beta(t)\) is via its Maximum Likelihood Estimator (MLE). However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][Sec.~3.2]{buenolarraz2021functional}.

It turns out that a natural alternative to the \(L^2\) model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that space. As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions. This approach has already been explored in \citet{berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use} for the case of functional classification.

Another set of techniques widely studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_i)\}\) of the process that better summarize it according to some optimality criterion. As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs}, but in general they have their own dedicated algorithms and procedures. Given the nature of our suggested Bayesian model, we can easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_i\}\), and thus provide a variable selection method along with the other estimators that naturally arise within our model. In this way, in addition to making predictions about the input data, we can evaluate exactly which marginals of the functional explanatory variable contain the most relevant information. These points-of-impact selection models for functional predictors have also appeared recently in the literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.

@@

As we can see, the covariance function (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS \incomment{¿Es necesario destacar alguna suposición/propiedad extra sobre \(K\)? Por ejemplo \(\int\int K^2 <\infty\) o que es simétrica y (semi-) definida positiva}. A linear operator \incomment{No sé muy bien qué calificativos ponerle: integral Hilbert-Schmidt operator, compact, self-adjoint, bounded, linear functional, ...}\ closely related to this covariance function is the so-called covariance operator, namely
\[
\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f \in L^2[0, 1].
\]
This operator is especially relevant due to the \textit{Karhunen-Loève} theorem~\cite[e.g.][Th.~1.5]{bosq2000linear} which asserts that a centered process \(X\) can be decomposed as
\begin{equation}\label{eq:karhunen-loeve}
X(t) = \sum_{i=1}^\infty \xi_i \phi_i(t),
\end{equation}
where \(\phi_i\) are the (orthonormal) eigenfunctions of its covariance operator \(\mathcal K\) and \(\xi_i = \int X\phi_i\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_i\), the eigenvalue associated with \(\phi_i\). This expansion is the main justification of many FDA techniques, such as functional principal component analysis (FPCA)~\cite[see for example][Ch.~8]{ramsay2005functional}.

@@

Despite the close connection between the process \(X\) and the space \(\mathcal H(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see][Cor.~7.1]{lukic2001stochastic}. As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined when \(x\) is a realization of \(X\). However, following Parzen's approach \citep[Th.~4E]{parzen1961approach} we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x^{-1}(f) := \Psi_X^{-1}(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \mathcal H(K)\). This notation often proves to be useful and convenient.

@@

In this section we present the precise models and Bayesian methodologies explored in this work. In the sequel, we will assume the notation from the previous sections and suppose without loss of generality that \(X\) is centered, i.e., \(m(t)=0\) for all \(t\in[0, 1]\). The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \mathcal H(K)\) and replacing the scalar product for \(\dotprod{x}{\alpha}_K\) in the \(L^2\) models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}. However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subset \(\mathcal H_0(K)\), i.e.,
\begin{equation}\label{eq:alpha-parameter-rkhs}
  \alpha(\cdot) = \sum_{j=1}^p \beta_j K(t_j, \cdot) \in \mathcal H_0(K).
\end{equation}

@@

The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the evidence of the available sample data. In view of expression~\eqref{eq:alpha-parameter-rkhs}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\mathcal H_0(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_j\}\) and the times \(\{t_j\}\) separately. Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x^{-1}(\alpha)\), where \(\Psi\) is Loève's isometry. Hence, taking into account that \(\Psi_X^{-1}(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we have
\[
\dotprod{x}{\alpha}_K \equiv \sum_{j=1}^p \beta_j x(t_j).
\]

@@

In any case, after selecting a suitable prior distribution \(\pi(\theta)\) for the parameter vector \(\theta\) on the chosen finite-dimensional parameter space, we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d. samples amounts to
\begin{equation}\label{eq:bayes-theorem}
  \pi(\theta \mid \mathcal D_n) \propto \left[ \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right]\pi(\theta).
\end{equation}
In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y\mid \theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.

A last technicality to discuss is the fact that in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant. Nevertheless, we can sample from the posterior distribution\footnote{Provided that the marginal likelihood \(\int \pi(Y\mid \theta)\pi(\theta)\, d\theta\) is finite.} using an algorithm of the family of MCMC sampling methods. These are subset of the more general Monte Carlo simulation methods, which are specifically designed to sample from a multi-dimensional distribution when only a function proportional to its density (or its log-density) is known. Generally speaking, these algorithms obtain samples from a continuous random variable by generating Markov chains whose stationary distribution is proportional to the function given. In short, they build stochastic processes of random walkers that travel the parameter space, selecting where to move based on certain criteria that make them more likely to generate samples from the intended distribution. The Metropolis-Hastings algorithm (e.g.~\cite{chib1995understanding}) is arguably the most well-known method in this regard.

@@

In the case of functional linear regression, the RKHS model considered (with \(\alpha\in \mathcal H_{0,p}(K)\)) is
\begin{equation}\label{eq:rkhs-model-linear}
  Y = \alpha_0 + \Psi^{-1}_X(\alpha) + \epsilon = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \epsilon,
\end{equation}
where \(\beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is an independent normally-distributed error term verifying \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\). When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and note that given the times \(\{t_j\}\) the model becomes a multivariate linear model with the \(X(t_j)\) as scalar covariates. Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information on \(\mathcal D_n\). For a generic element \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (\beta, \tau, \alpha_0, \sigma^2) \in \Theta_p\), the reinterpreted model assumes the form
\begin{equation}\label{eq:rkhs-model-linear-2}
  Y_i \mid X_i, \theta \ \stackrel{\text{i.i.d.}}{\sim} \mathcal N\left(\alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j), \ \sigma^2\right), \quad i =1,\dots, n.
\end{equation}

It is worth mentioning that the model remains linear in the sense that it fundamentally involves a random variable belonging to the linear span of the process \(X\) in \(L^2(\Omega)\). Furthermore, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model \eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}. In addition, as far as generalizations go the model is easily extensible, in the sense that several functional covariates could be added. In that case, as argued in \citet{grollemund2019bayesian} for a similar situation, if we were to set a prior distribution on all the parameters involved, we could recover the full posterior by looking alternately at the posterior distribution of each covariate conditional on the rest of them.

The prior distribution \(\pi(\theta)\) suggested for the parameter \(\theta \in \Theta_p\) is given by
\begin{align}\label{eq:prior-linear}
  \begin{split}
  \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
  \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
  \beta\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2(G_\tau + \eta I)^{-1}),
\end{split}
\end{align}
where \(I\) is the identity matrix, \(G_\tau = \mathcal X_\tau'\mathcal X_\tau\) is the Gram matrix corresponding to the matrix \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \eta \in \R^+\) and \(g\in \R\) are hyperparameters of the model. On the one hand, note the use of a joint prior distribution on \(\alpha_0\) and \(\sigma^2\), which is a widely used non-informative prior known as Jeffrey's prior \citep{jeffreys1946invariant}. In any event, the estimation of \(\alpha_0\) is straightforward, since \(\alpha_0 = \E[Y]\) by~\eqref{eq:rkhs-model-linear-2}, so we could have left this parameter out of the model altogether. On the other hand, the prior on \(\beta\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing} in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}.

@@

\begin{description}
  \item[Point estimates] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat \beta, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T(\{\theta^{(m)*}\}_m)\) and predict the responses in the usual way by
  \[
    \hat Y_i = \hat{\alpha}_0 + \sum_{j=1}^p \hat{\beta}_j X_i(\hat{t}_j), \quad i=1,\dots, n'.
  \]
  The choice of a specific summary statistic results in a veiled assumption of an underlying loss function between the estimated and real parameters. In general, the mean is more sensitive to outliers while the median is more robust, but the latter assumes a \(L^1\)-type loss function. On the other hand, the mode is also a good candidate because it represents the point of highest probability. At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
  \item[Distributional estimates] We can also look at the approximate posterior distribution as a whole and compute the distribution \(Y^*\) of the ``sampled responses''
  \[
  \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n', \ m=1,\dots,M\right\}
  \]
  following model~\eqref{eq:rkhs-model-linear-2}. Then, we can take the mean of all such samples as a proxy for the response:
  \[
  \hat Y_i = \frac{1}{M}\sum_{m=1}^M Y_i^{(m)*}, \quad i=1,\dots,n'.
  \]
  \item[Variable selection estimates] Lastly, we can focus only on the marginal posterior distribution of \(\tau\mid\mathcal D_n\) and perform variable selection via a point-estimate statistic \(T\). Specifically, we can consider the times \(\hat t_j = T(\{t_j^{(m)*}\}_m)\) and reduce the original data set to the scalar variables \(\{X_i(\hat t_j):\ i=1, \dots,n, \ j=1,\dots,p\}\). After this selection, we can treat the problem as a finite-dimensional regression task and use any of the well-known multivariate algorithms suited for this situation.
\end{description}

@@

As far as prior distributions go, we propose to use the same ones as we did in \eqref{eq:prior-linear} for the regression approach \incomment{Me gustaría probar con otras distribuciones a priori, especialmente para los coeficientes \(\beta\), ya que aquí no sé si tiene tanto sentido mantener \(\sigma^2\) como parámetro y usar el prior de Zellner. ¿Sabéis de alguna que se use en el caso de regresión logística?}. However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model~\eqref{eq:rkhs-model-logistic-2}.

@@

One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member or the parameter space and integrate it into the model. While in principle we could impose a prior distribution on \(p\) as well, we found that it would have some unwanted implications. For instance, it would make the implementation more complex, since the dimensionality of the parameters \(\beta\) and \(\tau\) would need to be fixed at a certain maximum value, but in the MCMC algorithm the working value of \(p\) would vary from one iteration to the next. In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time. Another possibility would be to adapt a purely Bayesian model selection technique \citep[see][Ch.~7]{piironen2017comparison, gelman1995bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models. In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g. BIC, AIC, \ldots) to select its optimal value.

As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments can be made:
\begin{itemize}
  \item For the expected value \(b_0\) we propose to use (an approximation of) the MLE of \(\beta\) (see Section~\ref{sec:results} for details on the implementation). Our numerical studies showed that the results were much better with this choice than, say, with a random or a null vector.
  \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation suggests that \(g=5\) is a good value.
  \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator. That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(G_\tau)\), where \(\lambda_{\max}(G_\tau)\) is the largest eigenvalue of the matrix \(G_\tau\) and \(\tilde\eta > 0\) is the actual tuning parameter. This standardization technique was used previously in the literature; see for example \citet{grollemund2019bayesian}.
\end{itemize}

>>>>

The most common linear regression model is the classical \(L^2\)-model, originally introduced in the first edition of the book by~\citet{ramsay2005functional} \incomment{fue realmente así?}. It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^d\) for that of the functional space \(L^2[0,1]\):
\begin{equation}\label{eq:l2-linear-model}
Y = \beta_0 + \dotprod{X}{\beta} + \epsilon = \beta_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
\end{equation}
\incomment{mejor poner el producto con el subíndice \(\langle \cdot , \cdot \rangle_2\)?}~where \(\beta_0\in \R\) and \(\epsilon\) is an independent error term with certain simplifying assumptions. In this case, the main parameter \(\beta=\beta(t)\) is a member of the infinite-dimensional space \(L^2[0, 1]\). Aside from the fact that functional (i.e. infinite-dimensional) optimization is a hard problem in and of itself \incomment{cita?}, there are a number of more subtle drawbacks that arise in this \(L^2\)-framework. For instance, \(L^2[0, 1]\) is a big space that contains many non-smooth or ill-behaved functions, but incidentally the model given by~\eqref{eq:l2-linear-model} does not include a finite-dimensional model based on a linear combination of projections of \(X\) as a particular case \citep[see][]{berrendero2019rkhs}. Moreover, the non-invertibility of the covariance operator associated with \(X\) (defined in Section~\ref{sec:rkhs}), which plays the role of the covariance matrix in the infinite case, invalidates the usual least squares theory.

A similar \(L^2\)-based functional logistic equation can be derived for the binary classification problem via the logistic function:
\begin{equation}\label{eq:l2-logistic-model}
  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\beta_0 - \dotprod{X}{\beta}\}},
\end{equation}
where \(\beta_0 \in \R\) and \(\beta \in L^2[0, 1]\). In this case, the most common way of estimating the slope function \(\beta(t)\) is via its Maximum Likelihood Estimator (MLE). However, not only do the same complications as in the linear regression model apply in this situation, but there is also the additional problem that in functional settings the MLE does not exist with probability one under fairly general conditions \citep[see][Sec.~3.2]{buenolarraz2021functional}.

It turns out that in both situations a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that space. As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions. These RKHS-based models have already been explored in \citet{berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018use} for the case of functional classification. The main contribution of this work is a parametric Bayesian approach to parameter estimation within these recently-proposed models, in which a prior distribution is imposed on the unknown function to  obtain a posterior distribution after seeing the data.

Another set of techniques widely studied in this context are variable selection methods, which aim to select the marginals \(\{X(t_i)\}\) of the process that better summarize it according to some optimality criterion. As it happens, some variable selection methods have already been proposed in the RKHS framework \citep[see for example][]{berrendero2019rkhs}, but in general they have their own dedicated algorithms and procedures. As will become apparent in the forthcoming sections, given the nature of our suggested Bayesian model we could easily isolate the marginal posterior distribution corresponding to a finite set of points \(\{t_i\}\), and thus provide a variable selection method along with the other estimators that naturally arise within our model. In this way, in addition to making predictions about the input data, we could evaluate exactly which marginals of the functional explanatory variable contain the most relevant information. These points-of-impact selection models for functional predictors have also appeared recently in the literature; see \citet{poss2020superconsistent}, \citet{berrendero2016variable} or \citet{ferraty2010most} by way of illustration.

@@

As we can see, the covariance function (sometimes referred to as the \textit{kernel}) plays a crucial role in characterizing the RKHS \incomment{¿Es necesario destacar alguna suposición/propiedad extra sobre \(K\)? Por ejemplo \(\int\int K^2 <\infty\) o que es simétrica y (semi-) definida positiva}. A linear operator \incomment{No sé muy bien qué calificativos ponerle: integral Hilbert-Schmidt operator, compact, self-adjoint, bounded, linear functional, ...}\ closely related to this covariance function is the so-called covariance operator, namely
\[
\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f \in L^2[0, 1].
\]
This operator is especially relevant due to the \textit{Karhunen-Loève} theorem~\cite[e.g.][Th.~1.5]{bosq2000linear}, which asserts that a centered process \(X\) can be decomposed as
\begin{equation}\label{eq:karhunen-loeve}
X(t) = \sum_{i=1}^\infty \xi_i \phi_i(t),
\end{equation}
where \(\phi_i\) are the (orthonormal) eigenfunctions of its covariance operator \(\mathcal K\) and \(\xi_i = \int X\phi_i\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_i\), the eigenvalue associated with \(\phi_i\). This expansion is the main justification for many FDA techniques, such as functional principal component analysis (FPCA)~\cite[see for example][Ch.~8]{ramsay2005functional}.

@@

Despite the close connection between the process \(X\) and the space \(\mathcal H(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}. As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined and lacks meaning when \(x\) is a realization of \(X\). However, following Parzen's approach \citep[Th.~4E]{parzen1961approach} we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x^{-1}(f) := \Psi_X^{-1}(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \mathcal H(K)\). This notation often proves to be useful and convenient.

@@

In this section we present the precise models and Bayesian methodologies explored in this work. In the sequel, we will assume the notation from the previous sections and suppose without loss of generality that \(X\) is centered, i.e., \(m(t)=0\) for all \(t\in[0, 1]\). The RKHS-based functional models under consideration are those obtained by taking a functional parameter \(\alpha \in \mathcal H(K)\) and replacing the scalar product for \(\dotprod{x}{\alpha}_K\) in the \(L^2\)-models~\eqref{eq:l2-linear-model} and~\eqref{eq:l2-logistic-model}. However, to further simplify things we will follow a parametric approach and suppose that \(\alpha\) is in fact a member of the dense subset \(\mathcal H_0(K)\), i.e.,
\begin{equation}\label{eq:alpha-parameter-rkhs}
  \alpha(\cdot) = \sum_{j=1}^p \beta_j K(t_j, \cdot) \in \mathcal H_0(K).
\end{equation}

@@

The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the evidence of the available sample data. In view of expression~\eqref{eq:alpha-parameter-rkhs}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\mathcal H_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\{\beta_j\}\) and the times \(\{t_j\}\) separately. Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x^{-1}(\alpha)\), where \(\Psi\) is Loève's isometry. Hence, taking into account that \(\Psi_X^{-1}(K(t, \cdot)) = X(t)\) by definition (see~\eqref{eq:loeves-isometry}), we have
\[
\dotprod{x}{\alpha}_K \equiv \sum_{j=1}^p \beta_j x(t_j).
\]

@@

In any case, after selecting a suitable prior distribution \(\pi(\theta)\) for the parameter vector \(\theta\) on the chosen finite-dimensional parameter space, we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d. samples amounts to
\begin{equation}\label{eq:bayes-theorem}
  \pi(\theta \mid \mathcal D_n) \propto \left[ \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right]\pi(\theta).
\end{equation}
In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y\mid X,\theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.

A last technicality to discuss is the fact that in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant. Nevertheless, we can sample from the posterior distribution\footnote{Provided that the marginal likelihood \(\int \pi(Y\mid \theta)\pi(\theta)\, d\theta\) is finite.} using an algorithm of the family of Markov Chain Monte Carlo (MCMC) sampling methods. These are subset of the more general Monte Carlo simulation methods, which are specifically designed to sample from a multi-dimensional distribution when only a function proportional to its density (or its log-density) is known. Generally speaking, these algorithms obtain samples from a continuous random variable by generating Markov chains whose stationary distribution is proportional to the given function. In short, they operate by building stochastic processes of random walkers that travel the parameter space, selecting where to move based on certain criteria that make them more likely to generate samples from the intended distribution. The Metropolis-Hastings algorithm \citep[e.g.][]{chib1995understanding} is arguably the most well-known method in this regard.

@@

In the case of functional linear regression, the RKHS model considered (with a functional parameter \(\alpha\in \mathcal H_{0,p}(K)\)) is
\begin{equation}\label{eq:rkhs-model-linear}
  Y = \alpha_0 + \Psi^{-1}_X(\alpha) + \epsilon = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \epsilon,
\end{equation}
where \(\beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is an independent normally-distributed error term verifying \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\). When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and note that given the times \(\{t_j\}\) the model becomes a multivariate linear model with the \(X(t_j)\) as scalar covariates. Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\mathcal D_n\). For a generic element \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (\beta, \tau, \alpha_0, \sigma^2) \in \Theta_p\), the reinterpreted model assumes the form
\begin{equation}\label{eq:rkhs-model-linear-2}
  Y_i \mid X_i, \theta \ \stackrel{\text{i.i.d.}}{\sim} \mathcal N\left(\alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j), \ \sigma^2\right), \quad i =1,\dots, n.
\end{equation}

It is worth mentioning that the model remains linear in the sense that it fundamentally involves a random variable belonging to the linear span of the process \(X\) in \(L^2(\Omega)\). Furthermore, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}. In addition, as far as generalizations go the model is easily extensible, seeing as how several functional covariates belonging to the RKHS could be added. In that case, as argued in \citet{grollemund2019bayesian} for a similar situation, if we were to set a prior distribution on all the parameters involved, we could recover the full posterior by looking alternately at the posterior distribution of each covariate conditional on the rest of them.

The prior distribution \(\pi(\theta)\) suggested for the parameter \(\theta \in \Theta_p\) is given by
\begin{align}\label{eq:prior-linear}
  \begin{split}
  \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
  \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
  \beta\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2(G_\tau + \eta I)^{-1}),
\end{split}
\end{align}
where \(I\) is the identity matrix, \(G_\tau = \mathcal X_\tau'\mathcal X_\tau\) is the Gram matrix corresponding to the matrix \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters of the model. On the one hand, note the use of a joint prior distribution on \(\alpha_0\) and \(\sigma^2\), which is a widely used non-informative prior known as Jeffrey's prior \citep{jeffreys1946invariant}. In any event, the estimation of \(\alpha_0\) is straightforward, since \(\alpha_0 = \E[Y]\) by~\eqref{eq:rkhs-model-linear-2}, so we could have left this parameter out of the model altogether. On the other hand, the prior on \(\beta\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing} in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}.

@@

\begin{description}
  \item[Point estimates] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat \beta, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T(\{\theta^{(m)*}\})\) and predict the responses in the usual way by
  \[
    \hat Y_i = \hat{\alpha}_0 + \sum_{j=1}^p \hat{\beta}_j X_i(\hat{t}_j), \quad i=1,\dots, n'.
  \]
  The choice of a specific summary statistic results in a veiled assumption of an underlying loss function between the estimated and real parameters. In general, the mean is more sensitive to outliers while the median is more robust, but the latter assumes a \(L^1\)-type loss function. On the other hand, the mode is also a good candidate because it represents the point of highest probability. At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
  \item[Distributional estimates] We can also look at the approximate posterior distribution as a whole and compute the distribution \(Y^*\) of the ``sampled responses''
  \[
  \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n', \ m=1,\dots,M\right\}
  \]
  following model~\eqref{eq:rkhs-model-linear-2}. Then, we can take the mean of all such samples as a proxy for the response:
  \[
  \hat Y_i = \frac{1}{M}\sum_{m=1}^M Y_i^{(m)*}, \quad i=1,\dots,n'.
  \]
  \item[Variable selection estimates] Lastly, we can focus only on the marginal posterior distribution of \(\tau\mid\mathcal D_n\) and perform variable selection via a point-estimate statistic \(T\). Specifically, we can consider the times \(\hat t_j = T(\{t_j^{(m)*}\})\) and reduce the original data set to the scalar variables \(\{X_i(\hat t_j):\ i=1, \dots,n, \ j=1,\dots,p\}\). After this selection, we can treat the problem as a finite-dimensional regression task and use any of the well-known multivariate algorithms suited for this situation.
\end{description}

@@

As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the regression approach \incomment{Me gustaría probar con otras distribuciones a priori, especialmente para los coeficientes \(\beta\), ya que aquí no sé si tiene tanto sentido mantener \(\sigma^2\) como parámetro y usar el prior de Zellner. ¿Sabéis de alguna que se use en el caso de regresión logística?}. However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model~\eqref{eq:rkhs-model-logistic-2}.

@@

One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member or the parameter space and integrate it into the model. While theoretically we could impose a prior distribution on \(p\) as well, we found that it would have some unwanted practical implications. For instance, it would make the implementation more complex, since the dimensionality of the parameters \(\beta\) and \(\tau\) would need to be fixed at a certain maximum value, but in the MCMC algorithm the working value of \(p\) would vary from one iteration to the next. In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time. Another possibility would be to adapt a purely Bayesian model selection technique \citep[see][Ch.~7]{piironen2017comparison, gelman1995bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models. In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g. BIC, DIC, \ldots) to select its optimal value \incomment{deberíamos elegir uno de los criterios?}.

As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments can be made:
\begin{itemize}
  \item For the expected value \(b_0\) we propose to use (an approximation of) the MLE of \(\beta\) (see Section~\ref{sec:results} for details on the implementation). Our numerical studies showed that the results were much better with this choice than, say, with a random or null vector.
  \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation suggests that \(g=5\) is a good value.
  \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator. That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(G_\tau)\), where \(\lambda_{\max}(G_\tau)\) is the largest eigenvalue of the matrix \(G_\tau\) and \(\tilde\eta > 0\) is the actual tuning parameter. This standardization technique was used previously in the literature; see for example \citet{grollemund2019bayesian}.
\end{itemize}
